---
title: "week-04-challenge"
author: "Sam Smith"
date: "2/8/2022"
output: html_document
---
STEP 1: Get your dictionary - make a function and pull in dictionaries
```{r}
load_dictionary <- function(filename){
  table <- read.table(filename,header=TRUE,sep="\t",stringsAsFactors = FALSE)
  return(table)
}
# load the dictionaries we will use and make them vectors
solution <- load_dictionary("google-10000-english-usa-no-swears.txt")
solution_list <- solution$words
valid <- load_dictionary("collins-scrabble-words-2019.txt")
valid_list <- valid$words
```

STEP 2: remove words from solution list so only contains the words found in valid list

```{r}
solution_list <- intersect(valid_list,solution_list) # there are now 8,336 elements in the list
```

STEP 3: Pick a random solution - remove non-5-letter words and pick a random one. Split the word into a vector
of letters.
```{r}
# x should be a vector
pick_solution <- function(x,word_length=5){
  # only keep words with 5 letters
  x <- x[which(nchar(solution_list)==word_length)]
  # randomly choose a word
  word <- sample(x,1,replace = FALSE)
  # make into a vector of characters
  word_vector <- strsplit(word,split=c())
  # strsplit returns a list so have to unlist it into a vector
  word_vector <- unlist(word_vector)
  return(word_vector)
}
solution <- pick_solution(solution_list,word_length=5)
```

STEP 4: Create the play wordle function which allows users to make guesses and returns information about their 
guesses.

```{r}
play_wordle <- function(answer,valid_list,num_guesses=6,word_length=5){
  # initialize the number of guesses the user has at 1
  x <- 1
  # tell the user the rules
  print(paste("You have",num_guesses,"chances to guess a", word_length, "letter word."))
  # show them the letters of the alphabet
  print(paste("Letters left:",paste(LETTERS,collapse=" ")))
  # prompt for guess
  guess <- readline("Enter your word here, then press <enter>: ")
  # make sure user inputs a guess of the right length
  while(nchar(guess)!=word_length){
    print(paste("Please guess a valid", word_length,"letter word."))
    guess <- readline("Enter a word here, then press <enter>: ")
  }
  # make sure user guesses a word in the valid list
  while(!toupper(guess) %in% valid_list){
    print(paste("Not a valid word."))
    guess <- readline("Enter a word here, then press <enter>: ")
  }
  # function that evaluates the guess that user inputs
  evaluate_guess <- function(guess,answer){
    # initialize vector to put feedback in
    feedback <- rep(NaN,word_length)
    guess_vector <- c()
    evaluation <- list("feedback"=feedback,"guess"=guess_vector)
    # split the guess into a vector of letters
    guess <- strsplit(guess,split=c())
    guess <- toupper(unlist(guess))
    # which spots in the word has the user found the right letter? Record those as *'s
    evaluation$feedback[which(guess == answer)] <- "*"
    # if the guess is completely right, return evaluation and guess
    if(setequal(evaluation$feedback,rep("*",5))==TRUE){
      evaluation$guess <- guess
    }
    # if the guess is not completely right, keep checking the answer
    else{
      # which spots in the word are incorrect and still need to be solved?
      unsolved_g <- which(!guess == answer)
      unsolved_a <- unsolved_g
      for(i in 1:length(unsolved_g)){
        # if a letter in an unsolved spot matches between guess and solution
        if(guess[unsolved_g[i]] %in% answer[unsolved_a]){
          # record that spot as a "+"
          evaluation$feedback[unsolved_g[i]] <- "+"
          # updated unsolved list for the answer
          unsolved_a <- unsolved_a[-c(which(guess[unsolved_g[i]] == answer[unsolved_a]))]
        }
        # letter in an unsolved spot in guess does not match solution
        else{
          evaluation$feedback[unsolved_g[i]] <- "-"
        }
      }
    evaluation$guess <- guess
    }
    # return feedback
    return(evaluation)
  }
  # evaluate the guess for the first time
  evaluation <- evaluate_guess(guess,solution)
  # going to start recording output that will be printed out when game ends
  eval_output <- evaluation$feedback
  guess_list <- evaluation$guess
  # if the answer is right, print out how many guesses it took and their feedback
  if(setequal(evaluation$feedback,rep("*",5))==TRUE){
    print("You guessed the word!")
    print(guess_list,row.names=FALSE)
    print("1/6")
    print(evaluation$feedback)
  }
  else{
    # print feedback
    print(evaluation$feedback)
    # record that they are moving on to guess 2
    x <- x + 1
    # reduce number of letters as an option based on the user's guess
    new_letters <- LETTERS[which(!LETTERS %in% evaluation$guess)]
    # keep allowing the user to guess until they get the right answer or they have guessed 6 times
    while(x < 7 && setequal(evaluation$feedback,rep("*",5)) == FALSE){
      #show users what letters they have left
      print(paste("Letters left:",paste(new_letters,collapse=" ")))
      # ask user for guess
      guess <- readline("Enter a word here, then press <enter>: ")
      # make sure user inputs a guess of the right length
      while(nchar(guess)!=word_length){
        print(paste("Please guess a", word_length,"letter word."))
        guess <- readline("Enter a word here, then press <enter>: ")
      }
      # make sure user guesses a word in the valid list
      while(!toupper(guess) %in% valid_list){
        print(paste("Not a valid word."))
        guess <- readline("Enter a word here, then press <enter>: ")
      }
      # evaluate the guess
      evaluation <- evaluate_guess(guess,solution)
      # add feedback to output that will be printed out when game ends
      eval_output <- rbind(eval_output,evaluation$feedback)
      # add guess to guess list to be printed at end of game
      guess_list <- rbind(guess_list,evaluation$guess)
      if(setequal(evaluation$feedback,rep("*",5))==TRUE){
        # print how many guesses it took for them to get the right answer
        print("You guessed the word!")
        print(guess_list,row.names=FALSE)
        print(paste0(x,"/6"))
        # print history of guesses
        rownames(eval_output) <- 1:x
        return(print(eval_output))
      }
      else{
        # print feedback
        print(evaluation$feedback)
        # reduce number of letters as an option based on user guess
        new_letters <- new_letters[which(!new_letters %in% evaluation$guess)]
        # record that user has made another guess
        x <- x + 1
      }
    }
    print("Sorry, you lose.")
    print(guess_list,row.names=FALSE)
    return(eval_output)
  }
}
```